---
title: "HW1_task2"
author: "Grace Bianchi"
date: "2023-02-02"
output: html_document
---

##Intro

This code wrangles and examines a small subset of seawater sample data from CalCOFI. Two linear regression models are used to predict oxygen saturation based on several chemical and physical variables, and compared using AIC, BIC, and cross validation. Specifically, one model examines the relationship between O2 saturation of seawater off California's coast and water temperature, salinity, and phosphate concentration; and the second model examines the relationship between oxygen saturation and water temperature, salinity, phosphate concentration, and depth.

**Data Summary from the CalCOFI site:** “Since 1949, hydrographic and biological data of the California Current System have been collected on CalCOFI cruises. The 70+ year hydrographic time-series includes temperature, salinity, oxygen and phosphate observations. In 1961, nutrient analysis expanded to include silicate, nitrate and nitrite; in 1973, chlorophyll was added; in 1984, C14 primary productivity incubations were added; and in 1993, CTD profiling began.”


**Data citation:** CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.  Accessed 1/10/2022.


### Data Setup and Wrangling

```{r setup, include=TRUE, message = FALSE, warning = FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(AICcmodavg)
library(janitor)
library(here)
library(broom)
library(kableExtra)
library(tidymodels)

```


```{r read in data}
# read in data
seawater_samples <- read_csv(here("calcofi_seawater_samples.csv"))

```

Complete Task 2 in a single well-organized .Rmd. Read in the seawater sample data (calcofi_seawaxter_samples.csv), then create and compare two multiple linear regression models:


### Multiple Linear Regression Model 1

Oxygen saturation is observed as a function of water temperature, salinity, and phosphate concentration.

```{r}
## multiple linear regression model 1 w Oxygen saturation as a function of water temperature, salinity, and phosphate concentration

f1 <- o2sat ~ t_deg_c + salinity + po4u_m
mdl1 <- lm(formula = f1, data = seawater_samples)

mdl1_summary <- summary(mdl1)

tidy(mdl1_summary) %>% 
  kable(caption = "Model 1 Variable Coefficients") %>% 
  kable_classic()

# AIC(mdl1)
# AIC: 618.3868
```


### Multiple Linear Regression Model 2

This model explores the relationship of oxygen saturation as a function of water temperature, salinity, phosphate concentration, and depth.

```{r}
## multiple linear regression model 2 w Oxygen saturation as a function of water temp, salinity, phosphate concentration, and depth.
f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m
mdl2 <- lm(formula = f2, data = seawater_samples)

mdl2_summary <- summary(mdl2)

tidy(mdl2_summary) %>% 
  kable(caption = "Model 2 Variable Coeffecients") %>% 
  kable_classic()

# AIC(mdl2)

# Residual standard error: 5.08 on 95 degrees of freedom
#Multiple R-squared:  0.9574,	Adjusted R-squared:  0.9557 
# F-statistic: 534.3 on 4 and 95 DF,  p-value: < 2.2e-16

# AIC: 615.7016
  
```

The adjusted R^2^ indicates that either model explains ~95% of the observed variance (**Model 1 Adjusted R^2^ =** `r mdl1_summary$adj.r.squared`,
**Model 2 Adjusted R^2^ =** `r mdl2_summary$adj.r.squared`). Next, AIC is used to identify the model that maximizes the likelihood of the parameter values given the data. 

## Identify better model using following 3 methods

### Method 1 - Akaike Information Criteria (AIC)


```{r AIC }
AIC(mdl1, mdl2) # comparing AIC values
#     df  AIC
# mdl1	5	618.3868		
# mdl2	6	615.7016

(aictab(list(mdl1, mdl2))) %>% kable(caption = "comparison of model using AIC") %>% 
  kable_classic() # putting values into table

#     K   AICc Delta_AICc AICcWt Cum.Wt      LL
# Mod2 6 616.60       0.00   0.77   0.77 -301.85
# Mod1 5 619.03       2.42   0.23   1.00 -304.19

```

After looking at the AIC values for both models, the second model is the "better" model due to the lower AIC score. The difference (delta = 2.42) between AIC values indicates a significant difference in model fit. 

### Method 2 - BIC

```{r}
bictab(list(mdl1, mdl2)) %>% kable() %>% kable_classic()

#      K    BIC Delta_BIC BICWt Cum.Wt      LL
# Mod2 6 631.33      0.00  0.51   0.51 -301.85
# Mod1 5 631.41      0.08  0.49   1.00 -304.19
```

BIC results for both model are too close to select the better option. 

### Method 3 - Ten-Fold Cross Validation

A ten-fold cross validation is performed on the two models, using root-mean-square error as the scoring method. 

```{r 10 fold CV}
set.seed(444) ### set seed for reproducibility! here to set the folds

### use a workflow that bundles the logistic model and a formula
seawater_lm <- linear_reg() %>% 
  set_engine("lm")

tidy_folds <- vfold_cv(seawater_samples, v = 10)

lm_tidy_wf1 <- workflow() %>% 
  add_model(seawater_lm) %>% 
  add_formula(f1)

lm_tidy_cv_f1 <- lm_tidy_wf1 %>% 
  fit_resamples(tidy_folds)


mdl_rmse <- collect_metrics(lm_tidy_cv_f1) %>% ### use functions from the tune package to extract metrics
  filter(.metric == "rmse") %>% 
  select(.metric, mean, n, std_err) %>% ## selecting data to make table with both model rmse
  rename(model = .metric) 

mdl_rmse[1:1] <- "mdl1" # pulling out rmse data

```

```{r}
# repeat for model 2
lm_tidy_wf2 <- workflow() %>% 
  add_model(seawater_lm) %>% 
  add_formula(f2)

lm_tidy_cv_f2 <- lm_tidy_wf2 %>% 
  fit_resamples(tidy_folds)

mdl2_rmse <- collect_metrics(lm_tidy_cv_f2) %>% ### use functions from the tune package to extract metrics
  filter(.metric == "rmse") %>% 
  select(.metric, mean, n, std_err) %>% 
  rename(model = .metric) 

mdl2_rmse[1:1] <- "mdl2" # pull out rmse data

mdl_rmse %>% # merge two tables
  full_join(mdl2_rmse) %>% 
  kable(caption = "Comparison of Model Root Mean Squared Error") %>% # create table
  kable_classic()

```

The root mean square error is lower for model 2, thus it is the better model. The model is trained using the full dataset to identify the coefficients for the final model.

# Final model
AIC and cross-validation both indicate Model 2 as the "best" model. There are several things to consider upon selecting a model; specifically the size of the data set and the number of variables. Since the data set is fairly small (n = 100), AIC is the better indicator model fit, coonsidering the extra parameter.

Our final model:
`r equatiomatic::extract_eq(mdl2, wrap = TRUE)`

and with coefficients in place:
`r equatiomatic::extract_eq(mdl2, wrap = TRUE, use_coefs = TRUE)`


```{r}
final_mdl <- lm(f2, data = seawater_samples)
summary(final_mdl) %>% 
  tidy() %>% 
  kable() %>%  kable_classic()
```

