---
title: "Task 1 & 2"
author: "Grace Bianchi"
date: "2023-01-25"
output: html_document
---

Info

Data Citation:  Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

```{r setup, include= TRUE, message = FALSE, warning = FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(patchwork)
library(readxl)
library(lubridate)
library(cowplot)

```

```{r}
sierra_amphibians_data <- read_xlsx(here("sierra_amphibians.xlsx")) %>% 
  mutate(survey_date = ymd(survey_date),
         year = year(survey_date))
  
```

A graph of total mountain yellow-legged frog (Rana muscosa) count each year across all water bodies, by life stage excluding the ‘EggMass’ level. In other words, you should find the total number of adult, subadult, and tadpole yellow-legged frogs observed in the entire study region by life stage and year, but you will not use the lake ID or amphibian_location in your analyses as additional grouping variables. Hint: Convert the date to a date with the {lubridate} package, then pull just the year using the lubridate::year() function...then you can group by year to get counts. 

```{r}
yellow_legged_frog <- sierra_amphibians_data %>% 
  filter(amphibian_species == "RAMU",
         amphibian_life_stage != "EggMass")

yellow_frog_summary <- yellow_legged_frog %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarise(n = n())

p1 <- ggplot(data = yellow_frog_summary, aes(x = year, y = n)) +
  geom_col(aes(fill = amphibian_life_stage)) +
  theme_minimal() +
  labs(x = "Year", y = "Number of Yellow-legged Frogs") +
  guides(fill=guide_legend(title="Life Stage"))

  
```

```{r}

yellow_frog_lake <- yellow_legged_frog %>% 
  filter(amphibian_life_stage != "Tadpole") %>% # remove tadpole observations
  group_by(lake_id) %>% 
  summarize(n = n()) %>% # find counts for each lake
  slice_max(n = 9, order_by = (n)) # 

# add Lake before each id number
yellow_frog_lake$lake_id <- sub("^", "Lake ", yellow_frog_lake$lake_id) 

p2 <- ggplot(data = yellow_frog_lake, aes(y = reorder(lake_id, n), x = n)) +
  geom_col(aes(fill = lake_id), show.legend = FALSE) +
  theme_minimal() +
  labs(x = "Total Observed Counts (n)",
       y = " ")
  
  
plot_grid(p1, p2)
```

