---
title: "Task 1 & 2"
author: "Grace Bianchi"
date: "2023-01-25"
output: html_document
---


**Background**

This code examines and wrangles amphibian data collected in the Sierra Lakes from 1995 to 2002. The mountain yellow-legged frog data is analyzed further to see the total number of observations during this study and the lakes with the most observations.

Data variable included in this analysis:
- lake_id
- amphibian_species: Rana muscosa (RAMU)
- amphibian_life_stage: Eggmass, tadpole, subadult, adult.
- amphibian_number: number observed


**Data Citation:**  Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

### Setup

```{r setup, include= TRUE, message = FALSE, warning = FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(patchwork)
library(readxl)
library(lubridate)
library(cowplot)
```


```{r read in the data}
sierra_amphibians_data <- read_xlsx(here("sierra_amphibians.xlsx")) %>% 
  mutate(survey_date = ymd(survey_date), 
         year = year(survey_date))
  
```

A graph of total mountain yellow-legged frog (Rana muscosa) count each year across all water bodies, by life stage excluding the ‘EggMass’ level. the total number of adult, subadult, and tadpole yellow-legged frogs observed in the entire study region by life stage and year 

### Total Mountain yellow-legged frog observations



```{r}
yellow_legged_frog <- sierra_amphibians_data %>% 
  filter(amphibian_species == "RAMU", # filter for yellow-legged frog
         amphibian_life_stage != "EggMass") # remove any eggmass observations 

yellow_frog_summary <- yellow_legged_frog %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarise(n = sum(amphibian_number))

p1 <- ggplot(data = yellow_frog_summary, aes(x = year, y = n, fill = )) +
  geom_col(aes(fill = amphibian_life_stage)) + 
  scale_fill_manual("legend", values = c("SubAdult" = "lightgreen", "Adult" = "orange", "Tadpole" = "cornflowerblue")) +
  theme_light() +
  labs(x = "Year", y = "Total Observed Counts (n)") + 
  guides(fill=guide_legend(title="Life Stage")) + # change legend title 
  theme(legend.position = c(0.1, 0.9))
  
```


```{r frog observations by lake}

yellow_frog_lake <- yellow_legged_frog %>% 
  filter(amphibian_life_stage != "Tadpole") %>% # remove tadpole observations
  group_by(lake_id) %>% 
  summarize(n = sum(amphibian_number)) %>% # find counts for each lake
  slice_max(n = 5, order_by = (n)) # select lakes with highest counts

# add Lake before each id number
yellow_frog_lake$lake_id <- sub("^", "Lake ", yellow_frog_lake$lake_id) 

p2 <- ggplot(data = yellow_frog_lake, aes(x = reorder(lake_id, n), y = n)) +
  geom_col(fill = "lightblue", show.legend = FALSE) +
  theme_light() +
  labs(y = " ",
       x = " ") 

```

```{r, fit.height = 6, fig.width = 7}
p1+ p2 +
  plot_annotation(title = 'Mountain yellow-legged frog observations', tag_levels = "A") & theme(plot.title = element_text(face = 'bold', size = 20))


```

**Figure 1:** A. Graph of the number of mountain yellow-legged frogs (Rana muscosa) observed since 1995 by life stage. B. Top 5 lakes with the highest number of observations of adult and subadult frogs from 1995 to 2002. 
